---
title: "FishCatches"
author: "Liam Schilling"
date: "6/16/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Data and Packages
```{r}
library(tidyverse)
library(RColorBrewer)
library(ggalt)
library(scales)
library(reshape2)
library(ggthemes)
df <- read.csv("fish_catches.csv")

# Loading personal ggplot theme

mytheme <- theme(panel.background = element_rect(fill = "#F0F0F0"),
                 plot.background = element_rect(fill = "#F0F0F0"),
                 panel.border = element_rect(colour="#f0f0f0", fill = NA),
                 panel.grid.major = element_line(colour="#d0d0d0",size=.75),
                 axis.text.y=element_text(size = 11,colour="#535353",face="bold"),
                 plot.title=element_text(face="bold", hjust=0,vjust=2,size = 20,colour="#3c3c3c"),
                 plot.caption = element_text(size=6),
                 axis.text.x = element_text(size = 11,colour="#535353",face="bold"),
                 plot.subtitle = element_text(size = 11, colour="#535353",face="bold", hjust=0,vjust=2),
                 legend.position = "none",
                 axis.title.y = element_text(size=11,colour="#535353",face="bold",lineheight = 1, vjust=-2),
                 axis.title.x = element_text(size=11,colour="#535353",face="bold",lineheight=1,vjust=0),
                 plot.margin = unit(c(1,1,.5,.7),"cm")
                 )
themeline <- geom_hline(yintercept=0,size=1.2,colour="#535353")
```

# Question 1: Which species are being fished more relative to history?

```{r}
dfh <- df %>% group_by(Species) %>% summarise("x14" = sum(X2014), "x06" = sum(X2006))

dfh <- dfh[order(-dfh$x06),]
x06top25 <- head(dfh, 25)

x06top25 <- melt(x06top25, id="Species")
dfh <- dfh[order(-dfh$x14),]
x14top24 <- head(dfh, 25)

ggplot(data = x06top25) + geom_dumbbell(aes(x = x06, xend = x14, y = Species),dot_guide = FALSE, size = 2, colour = "#bababa", colour_x = "#ffeb5b", colour_xend = "#cc0c02") + mytheme + labs(title = "Fish Catches (2006 vs. 2014)", y = "", x = "Amount Caught") + scale_x_continuous(limits = c(140000,8000000), breaks = seq(140000,8000000,750000), labels = comma)

ggplot(data = x06top25) + geom_bar(colour = "#535353", aes(x = reorder(Species,-value), y = value, fill = variable, alpha = .85), stat = "identity", position = "dodge") + mytheme + scale_fill_economist() + scale_y_continuous(limits = c(140000,8000000), breaks = seq(140000,8000000,750000), labels = comma, oob=rescale_none) + themeline
```

# Ideas

  + Create new data that shows change between each year 2014 - 2006, then make that graph
  
  
  + Create graph like the one above but insert the "average" species with the average change   they saw
```{r}
df$change <- df$X2014 - df$X2006
df2 <- df %>% group_by(Species) %>% summarise("Change" = sum(change))

df2 <- df2[-1,]
df2 <- df2[order(-df2$Change),]
df2$Species <- as.character(df2$Species)
avg <- mean(df2$Change)
df2 <- rbind(df2,"AVG" = c("AVG",avg))

df2$select <- 0
df2$select[df2$Species == "AVG"] <- 1

dffinal <- head(df2, 10)
dffinal <- rbind(dffinal, tail(df2, 11))

dffinal$Change <- as.integer(dffinal$Change)
dffinal$select <- as.factor(dffinal$select)

palette <- brewer.pal("YlGn", n = 3)
w <- ggplot(data = dffinal, aes(x = reorder(Species, -Change), y = Change, fill = select)) + geom_bar(stat = "identity", colour = "#535353", alpha = .95) + mytheme + themeline + labs(x = "Species", y = "Change (Difference between 2014 and 2006)\n", title = "Change In Fish Species Caught (2014 - 2006)", subtitle = "1.5 million less fish were caught in 2014 vs. 2016; however, not all species have experienced a decrease in fishing rates", caption = "Source: Kaggle.com") + scale_fill_manual(values = palette[2:3]) + scale_y_continuous(labels = comma, breaks = seq(-4000000,4000000,500000))

png(filename="FishChange.png", width = 2300, height = 1500, res = 200)
w
dev.off()
```

# it's clear that cod and mackeral got a pickup in fishing rates, what countries are the offenders?

```{r}
dfcountry <- df %>% group_by(Species, Country) %>% summarise(Total2014 = sum(X2014), Total2006 = sum(X2006)) %>% filter(Species == "COD" | Species == "MAC")

dfcountry$Change <- dfcountry$Total2014 - dfcountry$Total2006

dfcountry <- dfcountry[dfcountry$Change > 2000 | dfcountry$Change < -2000,]

ss <- ggplot(data = dfcountry, aes(x = reorder(Country,-Change), y = Change, fill = Species)) + geom_bar(stat = "identity", alpha = .92, colour = "#535353") + mytheme + themeline + scale_fill_manual(values = palette[2:3]) + theme(legend.position = c(.85,.96), legend.background = element_rect("#F0F0F0"), legend.direction="horizontal") + labs(x = "Country", y = "Change in Fish Caught\n", title = "Norway and Russia saw the biggest change in consumption\n") + guides(alpha = FALSE) + scale_y_continuous(breaks = seq(0,2000000,250000))

png(filename="countries.png", width = 2300, height = 1500, res = 200)
ss
dev.off()
```


# Make dumbell graph with change in species from 2006 to 2014, and include a species called "total"

```{r}
x06total <- as.integer(sum(df$X2006, na.rm=TRUE))
x14total <- as.integer(sum(df$X2014, na.rm=TRUE))
dfdumb <- df[,c(1,5,13)] %>% group_by(Species) %>% summarise(Total2014 = sum(X2014), Total2006 = sum(X2006))

dfdumb$Species <- as.character(dfdumb$Species)
dfdumb <- na.omit(dfdumb)
dfdumb <- rbind(dfdumb, "new" = c("Total",x14total,x06total))
dfdumb$Total2014 <- as.integer(dfdumb$Total2014)
dfdumb$Total2006 <- as.integer(dfdumb$Total2006)



dfdumb$change <- dfdumb$Total2014 - dfdumb$Total2006

dfdumb <- dfdumb[order(-dfdumb$change),]

dffinal <- head(dfdumb,10)
dffinal <- rbind(dffinal, tail(dfdumb,10))

a <- ggplot(data = dffinal) + geom_dumbbell(aes(y = reorder(Species,Total2006), x = Total2006, xend=Total2014), size = 1.5, colour = "#535353", colour_x = palette[2], colour_xend = palette[3], size_x = 3, size_xend = 3) + theme(legend.position = "top") + labs(title = "Fish Species With Greatest Change In Fishing Rates", subtitle = "While most species have seen a decrease in fishing rates, some species such as Cod and Mackeral\nhave seen increases over the past few years", caption = "Source: Kaggle.com",y = "Species", x = "Total Caught") + scale_x_continuous(labels = comma, breaks = seq(0, 30000000,5000000)) + mytheme + geom_text(data=data.frame(), aes(x=33000000,y = "Total",label = "2006"), hjust=0, size = 3.5, fontface="bold", colour = "#ADDD8E", nudge_y = .3) + geom_text(data=data.frame(), aes(x=31000000,y = "Total",label = "2014"), hjust=1.4, size = 3.5, fontface="bold", colour = "#31A354", nudge_y = .3)

png(filename="FishDumbbell.png", width = 2300, height = 1500, res = 200)
a
dev.off()
```

